---
title: "Shyshark Distribution"
author: "Emma Swanepoel"
date: "2024-02-23"
output: html_document
---

This is a GIS on shyshark distribution around southern Africa. 
Shysharks are endemic benthic species found around the southern African coastline. There are four species (*Haploblepharus edwardsii*, *H. pictus*, *H. fuscus*, and *H. kistnasamyi*). As the latter two are quite criptic, this visual representation will focus on the observations made of Happy Eddies or Puffadder shysharks (*H. edwardsii*) and Dark shysharks (*H. pictus*).

To run the code you will need the packages tidyverse, sp, sf, raster, lwgeom, terra, stars, exactextractr, hrbrthemes,cowplot, rinat, prettymapr, ggspatial, leaflet, mapview, htmltools, leafpop, rnaturalearth, and rnaturalearthdata. 

```{r cleaning the space, include=FALSE}
rm(list=ls())
```
```{r packages, eval=FALSE, include=FALSE}
install.packages(c("tidyverse", "sp", "raster", "sf", "lwgeom", "terra", "stars", "exactextractr"))
install.packages(c("cowplot", "hrbrthemes", "knitr", "leaflet", "htmltools", "rosm", "ggspatial", "rnaturalearth", "mapview", "tmap"))
install.packages("rmarkdown", "prettymapr", "leaflet", "htmltools", "rinat", "rnaturalearth", "rnaturalearthdata")
```
```{r load packages, include=FALSE}
my_packages <- c("tidyverse", "sp", "sf", "raster", "lwgeom", "terra", "stars", "exactextractr", "hrbrthemes","cowplot", "rinat", "prettymapr", "ggspatial", "leaflet", "mapview", "htmltools", "leafpop", "rnaturalearth", "rnaturalearthdata")
lapply(my_packages, require, character.only = TRUE)
```

```{r note, eval=FALSE, include=FALSE}
## Note: After going through this process once and using mapview to evaluate how many sharks were observed on land I have decided to remove brown shysharks from the analysis as most of the observations were "land sharks". All code including brown shyshark data has been removed from here on out. 
# Also just gonna call pbsouthwood out for a second because *almost all* of their observations of shysharks were on land. *frustrated noises* 
```

First we need to read in some shyshark data from iNaturalist. 

```{r reading in iNat data, echo=TRUE}
puffyshark <- get_inat_obs(taxon_name = "Haploblepharus edwardsii", maxresults = 600)
darkshark <- get_inat_obs(taxon_name = "Haploblepharus pictus", maxresults = 600)
```


Then we need to filter the observations to include only research grade observations and observations with a positional accuracy less than 100m, we also exclude faulty observations made in the Northern Hemisphere and all data with NA latitude values. 

```{r filtering data, echo=TRUE}
puffyshark <- puffyshark %>% filter(positional_accuracy < 100 & latitude <0 & !is.na(latitude) & quality_grade == "research")
darkshark <- darkshark %>% filter(positional_accuracy < 100 & latitude <0 & !is.na(latitude) & quality_grade == "research")
```

In order to actually plot these data we need to make the dataframes spatial objects of the class "sf". 

```{r transform into sf, echo=TRUE}
puffyshark <- st_as_sf(puffyshark, coords = c("longitude", "latitude"), crs = 4326)
darkshark <- st_as_sf(darkshark, coords = c("longitude", "latitude"), crs = 4326)
```

Very important!! We need to check the dataframes to make sure they've been transformed into spatial objects and check that the coordinate reference systems (CRS) are the same. 

```{r echo=TRUE}
class(puffyshark)
class(darkshark)
st_crs(puffyshark)
st_crs(darkshark)
```
Nice!! All systems go! 

Let's do some more filtering! We want to exclude observations of **sharks** made **on land** *(shoutout to ChatGPT for the help)*.
We start by creating a dataframe for the land surface.

```{r creating df for land, echo=TRUE}
world_land <- ne_countries(scale = "medium", returnclass = "sf", country = "south africa")
```

Now we want to spatially join this df to ID the observations on land for the sharks and create new dataframes *without those observations*. 

```{r remove landshark observations, echo=TRUE}
landsharks_ps <- st_join(puffyshark, world_land, join = st_disjoint) %>%
  filter(!is.na(iso_a2))
# Exclude land shark observations from puffyshark
oceanshark_ps <- puffyshark[puffyshark$geometry %in% landsharks_ps$geometry,]
# Spatial join to ID observations on land for darksharks
landsharks_ds <- st_join(darkshark, world_land, join = st_disjoint) %>%
  filter(!is.na(iso_a2))
# Exclude land shark observations from darkshark
oceanshark_ds <- darkshark[darkshark$geometry %in% landsharks_ds$geometry,]
```

Let's see if all that worked by plotting these data.

```{r plot, echo=TRUE}
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data = oceanshark_ps, color = "#85B22C")
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data = oceanshark_ds, color = "#422CB2")
```

Hmm. It would be more helpful if both species were on one map...

```{r overlay plot, echo=TRUE}
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data = oceanshark_ps, color = "#85B22C") + 
  geom_sf(data = oceanshark_ds, color = "#422CB2")
```
That's better!

Now the fun part - creating interacive maps with both species! 

```{r both spp leaflet, echo=TRUE}
leaflet() %>% 
  addTiles(group = "Default") %>%
  addCircleMarkers(data = oceanshark_ps, 
                   group = "Haploblepharus edwardsii", 
                   radius = 1, 
                   color = "#85B22C")%>%
  addCircleMarkers(data = oceanshark_ds, 
                   group = "Haploblepharus pictus", 
                   radius = 1, 
                   color = "#422CB2")%>% 
  addLegend(position = "topright", colors = c("#85B22C", "#422CB2"), labels = c("Puffadder shyshark", "Dark shyshark"))
```

Next we want to create an interactive map with mapview and create popups for each data point with links to the iNaturalist observations. 
So - first thing's first - let's make the links possible.

```{r iNat links for mapview, echo=TRUE}
loceanshark_ps <- oceanshark_ps %>%
  mutate(click_url = paste("<b><a href='", url, "'>Link to iNat observation</a></b>"))
loceanshark_ds <- oceanshark_ds %>%
  mutate(click_url = paste("<b><a href='", url, "'>Link to iNat observation</a></b>"))
```

Finally! Let's create an interactive map with links to each point's iNaturalist observation! 

```{r mapview of both spp with links, echo=TRUE}
mapview(oceanshark_ps, col.regions = "#85B22C", layer.name = "Puffadder shyshark", legend = TRUE,
        popup = 
          popupTable(loceanshark_ps,
                     zcol = c("user_login", "click_url", "id"))) + 
mapview(oceanshark_ds, col.regions = "#422CB2", layer.name = "Dark shyshark", legend = TRUE,
        popup = 
          popupTable(loceanshark_ds, 
                     zcol = c("user_login", "click_url", "id")))
```

~~ This RMarkdown was made with lots of crying and swearing by the author but hopefully it will be appreciated~~
