---
title: "Shyshark Distribution around Southern Africa"
author: "Emma Swanepoel"
date: "2024-02-21"
output: html_document
---

This is the distribution of shysharks around southern Africa. 
This map shows puffadder shyshark (*Haploblepharus edwardsii*) observations in green, brown shyshark (*Haploblepharus fuscus*) observations in blue, and dark shyshark (*Haploblepharus pictus*) observations in purple.

Note that you need to install and load these packages:
tidyverse, sp, sf, raster, lwgeom, terra, stars, exactextractr, hrbrthemes, rinat, prettymapr, ggspatial, leaflet, mapview, htmltools

~~Please excuse me while I butcher this until I figure it out~~


```{r call packages, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#install.packages(c("tidyverse", "sp", "raster", "sf", "lwgeom", "terra", "stars", "exactextractr"))
#install.packages("hrbrthemes", "rinat", "prettymapr")
```
```{r load packages, message=False, warning=FALSE, include=FALSE, paged.print=FALSE}
my_packages <- c("tidyverse", "sp", "sf", "raster", "lwgeom", "terra", "stars", "exactextractr", "hrbrthemes", "rinat", "prettymapr", "ggspatial", "leaflet", "mapview", "htmltools")
lapply(my_packages, require, character.only = TRUE)
```

### Reading in data from iNaturalist
```{r dataframes, message=TRUE, warning=FALSE, paged.print=FALSE}
puffyshark <- get_inat_obs(taxon_name = "Haploblepharus edwardsii", maxresults = 600)
brownshark <- get_inat_obs(taxon_name = "Haploblepharus fuscus", maxresults = 600)
darkshark <- get_inat_obs(taxon_name = "Haploblepharus pictus", maxresults = 600)
```

### Filtering observations to research grade
```{r filtering, message=TRUE, warning=FALSE, paged.print=FALSE}
puffyshark <- puffyshark %>% filter(positional_accuracy < 100 & latitude <0 & !is.na(latitude) & quality_grade == "research")
brownshark <- brownshark %>% filter(positional_accuracy < 100 & latitude <0 & !is.na(latitude) & quality_grade == "research")
darkshark <- darkshark %>% filter(positional_accuracy < 100 & latitude <0 & !is.na(latitude) & quality_grade == "research")
```

### Make the dataframes spatial objects of class = "sf"
```{r transform, message=TRUE, warning=FALSE, paged.print=FALSE}
puffyshark <- st_as_sf(puffyshark, coords = c("longitude", "latitude"), crs = 4326)
brownshark <- st_as_sf(brownshark, coords = c("longitude", "latitude"), crs = 4326)
darkshark <- st_as_sf(darkshark, coords = c("longitude", "latitude"), crs = 4326)
```

### Check the class
```{r class, message=TRUE, warning=FALSE, paged.print=FALSE}
class(puffyshark)
class(brownshark)
class(darkshark)
```


```{r plot individual, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data = puffyshark, color = "#85B22C")
ggplot() +
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data = brownshark, color = "#2C85B2")
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data = darkshark, color = "#422CB2")
```

### Let's try overlay and plot these data
```{r plot overlay, echo=TRUE, message=TRUE, warning=FALSE, paged.print=FALSE}
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data = puffyshark, color = "#85B22C") + 
  geom_sf(data = brownshark, color = "#2C85B2") + 
  geom_sf(data = darkshark, color = "#422CB2")
```

### Interactive maps
```{r leeaflet, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE, paged.print=FALSE}
leaflet() %>% 
  addTiles(group = "Default") %>%
  addCircleMarkers(data = puffyshark, 
                   group = "Haploblepharus edwardsii", 
                   radius = 1, 
                   color = "#85B22C")
leaflet() %>%
  addTiles(group = "Default") %>%
  addCircleMarkers(data = brownshark, 
                   group = "Haploblepharus fuscus", 
                   radius = 1,
                   color = "#2C85B2")
leaflet() %>%
  addTiles(group = "Default") %>%
  addCircleMarkers(data = darkshark, 
                   group = "Haploblepharus pictus", 
                   radius = 1, 
                   color = "#422CB2")
```

#### With layering
```{r leaflet overlay, echo=TRUE, message=TRUE, warning=FALSE, paged.print=FALSE}
leaflet() %>% 
  addTiles(group = "Default") %>%
  addCircleMarkers(data = puffyshark, 
                   group = "Haploblepharus edwardsii", 
                   radius = 1, 
                   color = "#85B22C")%>%
  addCircleMarkers(data = brownshark, 
                 group = "Haploblepharus fuscus", 
                 radius = 1,
                 color = "#2C85B2")%>%
  addCircleMarkers(data = darkshark, 
                   group = "Haploblepharus pictus", 
                   radius = 1, 
                   color = "#422CB2")
```

### Different interactive maps - not working
```{r mapview, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE, paged.print=FALSE}
#mapview(puffyshark, 
        #popup = 
         # popupTable(puffyshark, 
                    # zcol = c("user_login", "url")))
#mapview(brownshark, 
      #  popup = 
        #  popupTable(brownshark, 
                 #    zcol = c("user_login", "url")))
#mapview(darkshark, 
      #  popup = 
        #  popupTable(darkshark,
               #      zcol = c("user_login", "url")))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
